Quantifying the importance of an inference model in Bayesian phylogenetics

### 2020-05-29 pirouette_article.tex
1 hour of work

BEAST [\cite{drummond2007beast}] is a typical Bayesian phylogenetics tool, 
that needs both character data and priors to infer 
a posterior distribution of phylogenies.
Specifically, for the species tree prior - which describes 
the process of diversification - 
BEAST has built-in priors such as the Yule [\cite{yule}] and 
(constant-rate) birth-death [\cite{nee1994reconstructed}] models.
These simple tree priors are the most commonly used, as they represent some biologically realistic processes (i.e. branching at per species constant rates), while being computationally fast.
BEAST's successor, BEAST2 [\cite{bouckaert2019beast}],
has a package manager, that allows third-party users 
to extend existing functionalities.
For example, one can add novel diversification models 
by writing a BEAST2 package that contains the likelihood 
formula of a phylogeny under the novel diversification model, 
i.e. the prior probability of a species tree.
Many diversification models (and their associated probability algorithms) 
have been developed, e.g., models in which diversification is time-dependent [\cite{nee1994reconstructed,rabosky2008explosive}],
or diversity-dependent [\cite{etienne2012diversity}],
or where diversification rates change for specific lineages 
and their descendants [\cite{etienne2012conceptual, rabosky2014automatic, alfaro2009nine}].
Other models treat speciation as a process that takes 
time [\cite{rosindell2010protracted, etienne2012prolonging, lambert2015reconstructed}],
or where diversification rate
depends on one or more traits [\cite{maddison2007estimating, fitzjohn2012diversitree}].
Despite such a rich theoretical landscape, only a few of these 
diversification models are available as tree priors in BEAST2.

>>>>

BEAST [\cite{drummond2007beast}] is a typical Bayesian phylogenetics tool, 
that needs both character data and priors to infer 
a posterior distribution of phylogenies.
Specifically, for the species tree prior - which describes 
the process of diversification - 
BEAST has built-in priors such as the Yule [\cite{yule}] and 
(constant-rate) birth-death [\cite{nee1994reconstructed}] models.
These simple tree priors are the most commonly used, 
as they represent some biologically realistic processes (i.e. 
branching at per species constant rates), while being computationally fast.

\new{
  % 00002
  In this paper, we focus on BD models, instead of coalescent models [Explain
  these a bit],
  although these are just as important. 
}

\new{
  % 00001
  To allow users to extend the functionalities of BEAST
  by using plug-ins, BEAST2 was written [\cite{bouckaert2019beast}]
  and both tools are still developed independently.
}

For example, one can add novel diversification models 
by writing a BEAST2 package that contains the likelihood 
formula of a phylogeny under the novel diversification model, 
i.e. the prior probability of a species tree.
Many diversification models (and their associated probability algorithms) 
have been developed, e.g., models in which diversification is time-dependent [\cite{nee1994reconstructed,rabosky2008explosive}],
or diversity-dependent [\cite{etienne2012diversity}],
or where diversification rates change for specific lineages 
and their descendants [\cite{etienne2012conceptual, rabosky2014automatic, alfaro2009nine}].
Other models treat speciation as a process that takes 
time [\cite{rosindell2010protracted, etienne2012prolonging, lambert2015reconstructed}],
or where diversification rate
depends on one or more traits [\cite{maddison2007estimating, fitzjohn2012diversitree}].
Despite such a rich theoretical landscape, only a few of these 
diversification models are available as tree priors in BEAST2.
